---
title: "index.Rmd"
author: "Nathan"
date: "5/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
source("analysis.R")
```

# A-5 Data Report
This is a data report regarding shootings that occurred across the United
States in 2018.

## Summary Information
This summary information is analyzation of 2018 USA shooting that provide
relevant information calculated from the data set. There were a total of
`r num_shootings` shootings that occurred in the USA in 2018. In these
shootings, there were a tragic amount of `r total_lives_lost` total lives lost
in the year, with `r highest_deaths_in_day` deaths being the most to happen in
one day. `r highest_injuries_in_day` injuries was the highest amount of
injuries to occur in one day. The city which had the most impactful shooting
(most deaths to happen) was `r impacted_city`.

## Summary Table
The table below is an aggregate summary of information from the 2018 shootings
data. The table shows the top 10 states that had the most 2018 casualties in
descending order, while also providing the total amount of injured people in
that state (not in a particular order).
```{r table, echo = FALSE}
library(knitr)
kable(state_killed_data, col.names = c("State", "Total Casualties",
                                       "Total Injuries"), align = "c")
```

## Description of a Particular Incident
On `r date_of_incident`, one of the most deadly and tragic shootings occurred.
A teen gunman used an AR-15 style semi-automatic rifle to open fire on his
former high school located in `r location_of_incident`. The gunman, Nikolas
Cruz was found and arrested nearby in Coral Springs after murdering
`r number_killed` people, and injuring `r number_injured`. He impacted a total
of `r number_impacted` people with his weapon. Although Nikolas had no
accomplices, he still caused tragic damage.
[Link to article](https://www.nbcnews.com/news/us-news/police-respond-shooting-parkland-florida-high-school-n848101)

## Interactive Map
Below is an interactive map that is a data visualization to show the location
of each shooting that occurred in 2018. The sizes of the markers indicate
the impact of each shooting, the higher deaths/injuries mean the bigger
the markers. This lets us easily compare which locations shooting were more
or less severe.
```{r map, echo = FALSE}
library(leaflet)
interactive_map
```
One insight revealed by this map is that we can see that the east side of the
U.S. has far more shooting occurrences than the west cost. Another insight is
that we can see which areas has more impacted people based on how big the
markers are

## Plot of Choice
With the bar chart I've created below, I can answer the question, "Which month
in 2018 had the highest number of occurred shootings?" The bar chart is
organized by months in the x-axis, and the y-axis is the number of shootings
that happened in that month across the U.S.. I chose to use this visual
encoding because bar plots are best used to represent ordinal data, which is
what this data is.
```{r plot, echo = FALSE}
library(ggplot2)
plot_of_choice
```

One insight gained from this chart is we can easily see which months had higher 
occurrences of shootings. Another insight gained is that it seems most
shootings tend to happen in the middle of the year around summer time, as June,
July, and August are the top 3 months for amount of shootings.